Machine Learning Project 
========================================================

First we load the datasets.

```{r}
setwd('D:\\Data\\datasets')
trainingdata<-read.csv("pml-training.csv",header=TRUE,na.strings=c("NA",""))
testdata<-read.csv("pml-testing.csv",header=TRUE,na.strings=c("NA",""))

```

Now we look at the training data 
```{r}
str(trainingdata)
table(trainingdata$classe)
```

Removing variables with missing values (NAs)
```{r}

isNASum<-function(x){
  return (sum(is.na(x)))
}


bad_train<-apply(trainingdata,2,isNASum)==19216
train2<-trainingdata[,!bad_train]

bad_test<-apply(testdata,2,isNASum)==20
test2<-testdata[,!bad_test]

```
Remove unuseful variables, keep only numeric ones.
```{r}
removeVars<-grep("X|user_name|timestamp|new_window|num_window",names(train2))
train3<-(train2[,-c(removeVars,length(train2))])
test3<-(test2[,-c(removeVars,length(test2))])

```

check for near-zero covariates as well as covariates that are highly correlated. Remove them.
```{r}
library(caret)
nzv<-nearZeroVar(train3,saveMetrics=TRUE)
nzv
```

From above, it looks like there aren't any near-zero covariates. 

We'll look at correlations among covariates.


```{r fig.width=7, fig.height=8}
corrs<-cor(train3)
library(corrplot)
corrplot(corrs,method="square",tl.cex=0.5)

highCorr<-findCorrelation(corrs,cutoff=0.75)
train4<-cbind(classe=train2$classe,train3[,-highCorr])
test4<-test3[,-highCorr]

```

Split the training data into training/testing for model evaluation
```{r}
set.seed(1234)
inTrain <-createDataPartition(train4$classe,p=0.75,list=FALSE)
trainPart<-train4[inTrain,]
testPart<-train4[-inTrain,]
```

Random Forest algorithm to predict
```{r}
library(randomForest)
rfModel<-randomForest(classe~.,data=trainPart,importance=TRUE,ntree=500)
print(rfModel)
```
in-sample 
```{r}
out.test<-predict(rfModel,testPart)
table(testPart$classe, out.test)
```
out-sample 
```{r}
out.test<-predict(rfModel,test4)
answers<- as.vector(out.test[1:20])
#answers = rep("A", 20)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
